---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: false
---

This work will try to answer the questions what asked [here][1] one the page of **Peer Assignment 1** of the **"Reproducible Research"** course on Coursera. In this document I will show how did I find answers for that questions. Let's go ヽ(\^ᴗ\^)丿

## Loading and preprocessing the data

I will use the package with data what placed [here][2]. This data file was downloaded and placed inside the repository with work as file with name "*repdata-data-activity.zip*". Let's check if it exists and unzip it to extract the data. 

```{r}
unzip("repdata-data-activity.zip")
```

After that let's check if file with data was unpacked and load data into variable *activity*. File should has name "activity.csv" and i want to remove it after loading. 

```{r results='hide'}
activity = read.csv("activity.csv")
file.remove("activity.csv")
```

Let's check what's inside the variable.

```{r}
str(activity)
```

## What is mean total number of steps taken per day?

Now I need to calculate total number of steps per day. I will use **dplyr** package for that. Let's group our data by *date* and summarise *steps* quantity. 

```{r}
library(dplyr)
per_day_steps <- select(activity, date, steps)
per_day_steps <- group_by(per_day_steps, date)
per_day_steps <- summarise(per_day_steps, steps = sum(steps))
head(per_day_steps)
```

I will build histogram of the total number of steps taken each day using **ggplot2** package.

```{r}
library(ggplot2)
ggplot(per_day_steps, aes(x=steps)) + 
  geom_histogram() + 
  ggtitle("Total number of steps taken each day") + 
  xlab("Days") +
  ylab("Steps") 
```

Let's define mean and median of the total number of steps taken per day.

```{r results='asis'}
library(xtable)

report <- 
  data.frame(
    c(mean(per_day_steps$steps, na.rm = TRUE)),
    c(median(per_day_steps$steps, na.rm = TRUE))
  )

colnames(report) <- c("Mean", "Median")
rownames(report) <- c("steps")

print(xtable(report), type="html")
```

## What is the average daily activity pattern?

Firstly I will calculate data frame that will store mean of steps for each of intervals over all days using **dplyr** package. And after that I will plot the data using **ggplot2** package.

```{r}
activity_pattern <- select(activity, interval, steps) %>% 
  group_by(interval) %>% 
  summarise(steps = mean(steps, na.rm = TRUE))

ggplot(activity_pattern, aes(x=interval, y=steps)) +
  geom_line() +
  ggtitle("Average daily activity pattern") + 
  xlab("Steps") + 
  ylab("Time")
```

## Imputing missing values

Let't find the number of missed values in the original dataset. I will use **dplyr** package functions to find out these values.  

```{r}
na_count <- select(activity, steps) %>% filter(is.na(steps)) %>% count()
na_count[[1]]
```

It looks like count of NA's is `r na_count[[1]]`. We need to fix missed values somehow. 

I will use median for the interval throughout all other days that has the value for the interval. 

First of all I will find medians of intervals to use it as source of values.

```{r}
medians_of_intervals <- select(activity, interval, steps) %>% 
  group_by(interval) %>% 
  summarise(steps = median(steps, na.rm = TRUE))
  head(medians_of_intervals)
```

I will use the **apply()** with anonymous function to restore missed values in the cloned dataset. Let's do it. 

```{r}
restored_activity <- activity
restored_activity$steps <- as.numeric(
  apply(activity, 1, function(x) { 
    
      steps_value = x[1]
      interval_value = x[3]
      
      if (is.na(steps_value)){
   
        # Find value of median for this interval
        selected <- select(medians_of_intervals, steps, interval) %>% 
                      filter(interval == as.integer(interval_value))
      
        steps <- selected[[1]]  
        
      } else {
        
        steps <- steps_value
      }
      
      steps
    })
)
str(restored_activity)
```
  
That's nice. There are no more *NA* in our dataset. 

I will to recalculate total number of steps per day. But I will use fixed data this time.  

```{r}
fixed_per_day_steps <- select(restored_activity, date, steps) %>%
                  group_by(date) %>%
                  summarise(steps = sum(steps))
head(fixed_per_day_steps)
```

Now I will compare histograms of the total number of steps taken each day with and without *NA* values.

```{r fig.height=5, fig.width=10}
library(grid)
library(gridExtra)

plot1 <- ggplot(per_day_steps, aes(x=steps)) + 
  geom_histogram() + 
  ggtitle("Total number of steps taken each day\nwith NA values") + 
  xlab("Days") +
  ylab("Steps")

plot2 <- ggplot(fixed_per_day_steps, aes(x=steps)) + 
  geom_histogram() + 
  ggtitle("Total number of steps taken each day\nafter NA values were removed") + 
  xlab("Days") +
  ylab("Steps") 

grid.arrange(plot1, plot2, ncol = 2)
```

Means and Medians of these data also should be comapred.

```{r results='asis'}
report <- 
  data.frame(
    c(
      mean(per_day_steps$steps, na.rm = TRUE), 
      mean(fixed_per_day_steps$steps, na.rm = TRUE)
    ),
    c(
      median(per_day_steps$steps, na.rm = TRUE),
      median(fixed_per_day_steps$steps, na.rm = TRUE)
    )
  )

colnames(report) <- c("Mean", "Median")
rownames(report) <- c("steps (with NA)", "steps (without NA)")

print(xtable(report), type="html")
```


It looks like average count of steps per day was reduced after we've imputted missing values. 

## Are there differences in activity patterns between weekdays and weekends?




[1]:https://class.coursera.org/repdata-015/human_grading/view/courses/973516/assessments/3/submissions
[2]:https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip